\documentclass[twoside]{article}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{subfigure}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{ulem}

\newcommand{\myheader}[1]{\textcolor{black}{\textbf{#1}}}
\setlength{\parindent}{0in}

\begin{document}
	
	{\bf Probability and Statistics} \hfill {\bf
		ARE 210 : Section Syllabus}
	
	\vspace{0.5em}
	
	\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll@{\extracolsep{\fill}}l}
		\hline
		\multicolumn{4}{c}{} \\ [-0.5em]
		{\bf Professor} & Aprajit Mahajan & {\bf GSI} & John Loeser \\
		{\bf GSI OH 1} & M 1-2, or by appt. & {\bf GSI OH 1 location} & Giannini 236 \\
		{\bf GSI OH 2} & T 8-9, or by appt. & {\bf GSI OH 2 location} & Giannini 234 \\
		{\bf Section time} & T 9-10 & {\bf Section location} & Giannini 234 \\
		& & {\bf email} & \href{mailto:john.a.loeser@berkeley.edu}{john.a.loeser@berkeley.edu} \\
		\multicolumn{4}{c}{} \\ [-0.5em]
		\hline
	\end{tabular*}
	\vspace{10pt} \bigskip
	
	The purpose of section (once weekly, 1 hour) is to reinforce the material from the course. I'll spend the first 20 minutes of each section reviewing material from lecture, and we'll spend the remainder working through practice problems. \\
	
	The section syllabus and notes will all be available at \href{http://www.github.com/johnloeser/are210}{www.github.com/johnloeser/are210}.
	
	\vspace{10pt}
	\begin{tabular}{ll}
		\bf{Aug 29}  &  Introduction \\
		\bf{Sep 5, 12, 19, 26} &  TBA \\
		\bf{Oct 3} &  No section \\
		\bf{Oct 10, 17, 24, 31} &  TBA \\
		\bf{Nov 7, 14, 21, 28} & TBA \\
		\bf{Dec 5} & TBA \\
	\end{tabular}
	%\newpage
	\bigskip
	
	{\bf Attendance}: Attendance isn't mandatory, but section should be useful so you should attend. If it isn't useful, let me know and we'll make it useful (see Feedback). \\
	
	{\bf E-mail policy}: Please begin the subject with [are210]. I'll do my best to respond to all emails within a business day. If I don't respond within two business days, your email was probably buried, so sending me a quick reminder to reply would be helpful. \\
	
	{\bf Office hours}: Office hours (twice weekly, 1 hour) are for discussing anything related to course material, very loosely defined. It's a good opportunity to get more specific, personalized feedback. No office hours on holidays (e.g. Labor Day) and likely none the week of Oct 2. If office hours won't be held a given Mon/Tue I'll make an announcement on bCourses the week before. If you can't make one of the two office hours, send me an email and we can find another time to meet. \\
	
	{\bf Feedback}: If you don't find section useful or would prefer something change, please let me know immediately and we can work to change it. You can leave anonymized feedback at \href{https://goo.gl/forms/iGAxx72YI3zYnkEI2}{https://goo.gl/forms/iGAxx72YI3zYnkEI2}. If you want to fix something specific in the section notes, bonus points for submitting a pull request. \\
	
	{\bf Special accommodations}: If you require any disability-related accommodations during our sections, lectures, or exams, please inform Aprajit as soon as possible. Generally, you have until the second week of class. For such accommodations, you must provide a supporting letter from the Disabled Students' Program (DSP), whose website is \href{http://dsp.berkeley.edu}{dsp.berkeley.edu}. \\
	
	{\bf Honor code}: As a member of the UC Berkeley community, I act with honesty, integrity, and respect for others.
	
\end{document}